---
title: "Client report"
output:
  pdf_document: default
  html_document:
    toc: true
    number_sections: true
    df_print: paged
params:
  dataset: "auto"   # "auto" OR a path like "data/raw/week4_dataset.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Find the project root robustly (anchor on this Rmd's location)
find_project_root <- function() {
  # Where is the current Rmd file?
  rmd <- knitr::current_input(dir = TRUE)

  # If knitr can't report it for some reason, fall back to this file’s folder assumption
  if (is.null(rmd) || !nzchar(rmd)) {
    # Assume we are in reports/ and root is one level up
    return(normalizePath(".."))
  }

  # Start search from the folder containing the Rmd
  start <- normalizePath(dirname(rmd))

  # Prefer a definitive marker file in your repo:
  markers <- c("st422-week4.Rproj", "DATA_DICTIONARY.md")
  for (m in markers) {
    # Walk upwards until we find the marker
    here <- start
    for (i in 1:10) {
      if (file.exists(file.path(here, m))) return(here)
      parent <- normalizePath(file.path(here, ".."))
      if (parent == here) break
      here <- parent
    }
  }

  # Last-resort: assume the Rmd sits in <root>/reports
  normalizePath(file.path(start, ".."))
}

ROOT <- find_project_root()

# Make both knitr and base::source resolve relative paths from the project root
knitr::opts_knit$set(root.dir = ROOT)
setwd(ROOT)

# Defensive check with a clear error message
if (!file.exists("src/01_load_data.R")) {
  stop("Cannot find src/01_load_data.R from project root: ", ROOT,
       "\nCheck you opened the .Rproj and that the repo structure is intact.")
}

source("src/01_load_data.R")
source("src/02_table1.R")
source("src/03_figures.R")
```

# Introduction

This brief report assesses whether the dataset supports a clear and defensible message that could be communicated to a general audience. The focus is on a small number of interpretable comparisons, with explicit limitations and a practical next step.

# Client brief

> We would like you to assess whether our dataset supports a clear and defensible message that could be communicated to a general audience. In practical terms, we want to know whether the data show a pattern that is relevant to our decision-making, sufficiently consistent to justify a public-facing narrative, and unlikely to be explained by obvious artefacts such as differences in the make-up of the groups, missing or incomplete data, or changes in definitions over time.
>
> Please provide an initial evidence-led summary that (i) states the most defensible claim the data can support at this stage, (ii) sets out the key evidence for that claim, (iii) explains the main uncertainties and limitations and their likely impact, and (iv) recommends the most useful next step to strengthen confidence in the story.

# Data description

```{r load}
loaded <- load_client_data(params$dataset)
df <- loaded$df
meta <- loaded$meta
```

The data file used was: `r meta$source_path`.

```{r time_series_note, results='asis'}
if (isTRUE(meta$is_time_series)) {
  cat(paste0("**Note:** this file contains repeated monthly observations. ",
             "To keep the deliverable structure (one record per customer), ",
             "this report analyses the most recent month in the file: **", meta$snapshot_month, "**. ",
             "A month-by-month stability check is recommended in the final section.\n\n"))
}
```

Each record in this report represents a single customer in the analysed extract. Customers are split into two groups (Control and Treatment) intended to represent whether the customer received a marketing intervention.

# Data description

This report uses the field definitions provided by the client (see `DATA_DICTIONARY.md`). Where definitions are provisional, we flag the most decision-relevant uncertainties in the limitations section.

# Summary of data characteristics

```{r checks, results='asis'}
n_total <- nrow(df)
n_groups <- df %>% count(group)
missing_purchase <- mean(is.na(df$made_purchase_30d))

cat(paste0("- Records analysed: ", n_total, "\n"))
cat(paste0("- Group counts: ", paste0(n_groups$group, "=", n_groups$n, collapse = ", "), "\n"))
cat(paste0("- Missing outcome (made_purchase_30d): ", scales::percent(missing_purchase, accuracy = 0.1), "\n"))
```

## Table 1: baseline characteristics by group

```{r table1}
tab1 <- make_table1(df)
write_table1(tab1, "outputs/tables/table1.csv")
knitr::kable(tab1, caption = "Table 1. Baseline summary by group.")
```
This is a Table 1, wow!

**Interpretation.** Table 1 summarises how similar the Control and Treatment groups look on key observable characteristics. Large differences here can explain apparent outcome differences without requiring a true intervention effect.

# Claim

```{r claim, results='asis'}
rates <- df %>%
  filter(!is.na(made_purchase_30d)) %>%
  group_by(group) %>%
  summarise(p = mean(made_purchase_30d == "Yes"), n = n(), .groups = "drop")

p_t <- rates$p[rates$group == "Treatment"]
p_c <- rates$p[rates$group == "Control"]
diff_pp <- 100*(p_t - p_c)

if (length(diff_pp) == 1 && is.finite(diff_pp)) {
  if (diff_pp > 0) {
    cat(sprintf("In this extract, the Treatment group shows a higher 30-day purchase rate than Control (difference: %.1f percentage points).", diff_pp))
  } else if (diff_pp < 0) {
    cat(sprintf("In this extract, the Treatment group shows a lower 30-day purchase rate than Control (difference: %.1f percentage points).", abs(diff_pp)))
  } else {
    cat("In this extract, the Treatment and Control groups have the same 30-day purchase rate.")
  }
} else {
  cat("In this extract, the 30-day purchase rate comparison could not be computed due to missing data.")
}
```

# Data and approach

We compare outcomes between Treatment and Control using simple descriptive summaries. The goal is not to prove causality, but to identify whether there is a pattern that is large enough and sufficiently credible to justify a cautious public-facing narrative, and to identify the single most useful next step to strengthen confidence.

# Evidence

## Visualisation 1: purchase rate by group

```{r fig1}
p1 <- make_purchase_rate_plot(df)
save_plot(p1, "outputs/figures/fig1_purchase_rate.png")
p1
```

**Interpretation.** This chart shows the percentage of customers who purchased within 30 days, for Treatment vs Control. The error bars reflect uncertainty from the sample size and help indicate whether the difference is likely to be due to random variation.

## Visualisation 2: average order value by group

```{r fig2, results='asis'}
p2 <- make_aov_plot(df)
if (!is.null(p2)) {
  save_plot(p2, "outputs/figures/fig2_aov_boxplot.png")
  print(p2)
} else {
  cat("Average order value is not available in this dataset.")
}
```

**Interpretation.** This chart summarises whether one group tends to have higher typical spend than the other. Differences here can matter for interpretation, because higher-value customers may respond differently regardless of the intervention.

# Uncertainty and limitations

1.  **Group differences may reflect who was selected, not the intervention.** If Treatment and Control differ in key baseline characteristics (Table 1), then outcome differences may be explained by selection rather than the intervention itself. This would weaken the credibility of a public-facing claim of impact.

2.  **Outcome timing and definitions are provisional.** The definitions (for example how "30 days" is anchored, and whether purchases are counted when placed or fulfilled) could change the measured purchase rate. If the definition shifted, the story could change materially.

3.  **If the source file is time-series (Week 5), this report uses only the most recent month.** The claim may not hold consistently over time. If seasonality or a recording change is present, a single-month snapshot may be misleading.

# Recommendation and next steps

The most useful next step is to confirm the intervention definition and eligibility rules (including anchor dates), then repeat the comparison with one simple robustness check: stratify the purchase rate by **customer_type** (New vs Existing) to see whether the pattern holds within comparable customers. If this is a time-series dataset, extend the analysis to month-by-month purchase rates to check stability across time and rule out seasonal or definition-driven artefacts.
