---
title: "Client report"
output:
  pdf_document: default
  html_document:
    toc: true
    number_sections: true
    df_print: paged
params:
  dataset: "auto"   # "auto" OR a path like "data/raw/week4_dataset.csv"
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Find the project root robustly (anchor on this Rmd's location)
find_project_root <- function() {
  # Where is the current Rmd file?
  rmd <- knitr::current_input(dir = TRUE)

  # If knitr can't report it for some reason, fall back to this file’s folder assumption
  if (is.null(rmd) || !nzchar(rmd)) {
    # Assume we are in reports/ and root is one level up
    return(normalizePath(".."))
  }

  # Start search from the folder containing the Rmd
  start <- normalizePath(dirname(rmd))

  # Prefer a definitive marker file in your repo:
  markers <- c("st422-week4.Rproj", "DATA_DICTIONARY.md")
  for (m in markers) {
    # Walk upwards until we find the marker
    here <- start
    for (i in 1:10) {
      if (file.exists(file.path(here, m))) return(here)
      parent <- normalizePath(file.path(here, ".."))
      if (parent == here) break
      here <- parent
    }
  }

  # Last-resort: assume the Rmd sits in <root>/reports
  normalizePath(file.path(start, ".."))
}

ROOT <- find_project_root()

# Make both knitr and base::source resolve relative paths from the project root
knitr::opts_knit$set(root.dir = ROOT)
setwd(ROOT)

# Defensive check with a clear error message
if (!file.exists("src/01_load_data.R")) {
  stop("Cannot find src/01_load_data.R from project root: ", ROOT,
       "\nCheck you opened the .Rproj and that the repo structure is intact.")
}

source("src/01_load_data.R")
source("src/02_table1.R")
source("src/03_figures.R")
```


# Introduction

This brief report assesses whether the dataset supports a clear and defensible
message that could be communicated to a general audience. The focus is on a
small number of interpretable comparisons, with explicit limitations and a
practical next step.

# Client brief

> We would like you to assess whether our dataset supports a clear and defensible
  message that could be communicated to a general audience. In practical terms,
  we want to know whether the data show a pattern that is relevant to our
  decision-making, sufficiently consistent to justify a public-facing narrative,
  and unlikely to be explained by obvious artefacts such as differences in the
  make-up of the groups, missing or incomplete data, or changes in definitions
  over time.
>
> Please provide an initial evidence-led summary that (i) states the most
  defensible claim the data can support at this stage, (ii) sets out the key
  evidence for that claim, (iii) explains the main uncertainties and limitations
  and their likely impact, and (iv) recommends the most useful next step to
  strengthen confidence in the story.

# Data description

```{r load}
loaded <- load_client_data(params$dataset)
df <- loaded$df
meta <- loaded$meta
```

The data file used was: `r meta$source_path`.

```{r time_series_note, results='asis'}
if (isTRUE(meta$is_time_series)) {
  cat(paste0("**Note:** this file contains repeated monthly observations. ",
             "To keep the deliverable structure (one record per customer), ",
             "this report analyses the most recent month in the file: **", meta$snapshot_month, "**. ",
             "A month-by-month stability check is recommended in the final section.\n\n"))
}
```

Each record in this report represents a single customer in the analysed extract.
Customers are split into two groups (Control and Treatment) intended to
represent whether the customer received a marketing intervention.

# Data description

This report uses the field definitions provided by the client (see 
`DATA_DICTIONARY.md`). Where definitions are provisional, we flag the most 
decision-relevant uncertainties in the limitations section.

# Summary of data characteristics

```{r checks, results='asis'}
n_total <- nrow(df)
n_groups <- df %>% count(group)
missing_purchase <- mean(is.na(df$made_purchase_30d))

cat(paste0("- Records analysed: ", n_total, "\n"))
cat(paste0("- Group counts: ", paste0(n_groups$group, "=", n_groups$n, collapse = ", "), "\n"))
cat(paste0("- Missing outcome (made_purchase_30d): ", scales::percent(missing_purchase, accuracy = 0.1), "\n"))
```

## Table 1: baseline characteristics by group

```{r table1}
tab1 <- make_table1(df)
write_table1(tab1, "outputs/tables/table1.csv")
knitr::kable(tab1, caption = "Table 1. Baseline summary by group.")
```

**Interpretation.** Table 1 summarises how similar the Control and Treatment
groups look on key observable characteristics. Large differences here can
explain apparent outcome differences without requiring a true intervention
effect.

# Claim

```{r claim, results='asis'}
rates <- df %>%
  filter(!is.na(made_purchase_30d)) %>%
  group_by(group) %>%
  summarise(p = mean(made_purchase_30d == "Yes"), n = n(), .groups = "drop")

p_t <- rates$p[rates$group == "Treatment"]
p_c <- rates$p[rates$group == "Control"]
diff_pp <- 100*(p_t - p_c)

if (length(diff_pp) == 1 && is.finite(diff_pp)) {
  if (diff_pp > 0) {
    cat(sprintf("Evidence suggests the intervention had a positive effect, with the Treatment group showing a %.1f percentage point higher 30-day purchase rate than Control.", diff_pp))
  } else if (diff_pp < 0) {
    cat(sprintf("Evidence suggests the intervention had a negative effect, with the Treatment group showing a %.1f percentage point lower 30-day purchase rate than Control.", abs(diff_pp)))
  } else {
    cat("Evidence suggests the intervention had no effect, with the Treatment and Control groups showing the same 30-day purchase rate.")
  }
} else {
  cat("The 30-day purchase rate comparison could not be computed due to missing data.")
}
```

**{Yueran} Refined one-sentence claim:** The Treatment group demonstrates a meaningful difference in 30-day purchase behavior compared to Control, though this pattern requires time-series validation to confirm stability across multiple periods and rule out seasonal or recording artifacts.

# Data and approach

We compare outcomes between Treatment and Control using simple descriptive
summaries. The goal is not to prove causality, but to identify whether there is
a pattern that is large enough and sufficiently credible to justify a cautious
public-facing narrative, and to identify the single most useful next step to
strengthen confidence.

# Evidence

## Visualisation 1: purchase rate by group

```{r fig1}
p1 <- make_purchase_rate_plot(df)
save_plot(p1, "outputs/figures/fig1_purchase_rate.png")
p1
```

**Interpretation.** This chart shows the percentage of customers who purchased
within 30 days. The gap between bars represents the estimated impact of the
intervention. If the error bars do not overlap substantially, we can be more
confident that this difference is a real effect rather than random chance.

**{Yueran} Plain-language interpretation for client:** This figure directly supports the claim by showing whether the marketing intervention influenced customer purchasing decisions. A larger bar for Treatment means more customers bought something after receiving the intervention, which indicates the campaign may be working to drive conversions.

## Visualisation 2: average order value by group

```{r fig2, results='asis'}
p2 <- make_aov_plot(df)
if (!is.null(p2)) {
  save_plot(p2, "outputs/figures/fig2_aov_boxplot.png")
  print(p2)
} else {
  cat("Average order value is not available in this dataset.")
}
```

**Interpretation.** This chart compares the typical spending amount per order.
This helps checks if we are driving volume but losing value (e.g., if the
intervention attracted only low-value purchases). Ideally, we want to see similar
or higher order values in the Treatment group.

**{Yueran} Plain-language interpretation for client:** This figure helps assess whether the intervention brings in quality customers or just bargain hunters. If Treatment customers spend similar or more per order than Control, it means the campaign is attracting valuable buyers, not just increasing low-value transactions.

# Uncertainty and limitations

1. **Group differences may reflect who was selected, not the intervention.** 
   If Treatment and Control differ in key baseline characteristics (Table 1), 
   then outcome differences may be explained by selection rather than the 
   intervention itself. This would weaken the credibility of a public-facing 
   claim of impact.

2. **Outcome timing and definitions are provisional.** The definitions (for 
   example how "30 days" is anchored, and whether purchases are counted when 
   placed or fulfilled) could change the measured purchase rate. If the 
   definition shifted, the story could change materially.

3. **If the source file is time-series (Week 5), this report uses only the most recent month.** 
   The claim may not hold consistently over time. If seasonality or a recording 
   change is present, a single-month snapshot may be misleading.

# Recommendation and next steps

The most useful next step is to confirm the intervention definition and
eligibility rules (including anchor dates), then repeat the comparison with one
simple robustness check: stratify the purchase rate by **customer_type** (New vs
Existing) to see whether the pattern holds within comparable customers. If this
is a time-series dataset, extend the analysis to month-by-month purchase rates
to check stability across time and rule out seasonal or definition-driven
artefacts.
